Openness and computational reproducibility in plant pathology: where we stand and a way forward

### 2023-01-05 analysis/paper/paper.Rmd
Fix grammatical mistakes

However, in the biological sciences it is not always possible to use identical test material or environmental conditions, e.g., field trials that span years and locations or complex glasshouse experiments.
In other cases there is insufficient time and resources to reproduce the entire study.
Therefore, we will follow Peng's [-@Peng2009] definition that provides clear guidelines for a minimum standard of "reproducible research":

@@

The criteria scores were checked for inconsistencies using the evaluators' comments in the spreadsheet and based on the software used.
In some cases these comments were used to adjust the criteria score to align with the scoring definitions.
This was not possible in all cases but only in cases where the evaluator had left comments that could be checked and acted upon.
For example, a reviewer noted that all data were deposited in GenBank but the data score was '0', this was changed to '3' to align with the definition.
Or in other cases, some software packages used were Excel add-ins, but Excel was not recorded as being used; Excel was added to the list of software used to rectify this.

@@

Much of the software that was used in the papers examined were Free Open-Source Software (FOSS) packages, which means that the workflow can easily be recreated by anyone with the proper skills in using the software.
However, the top three most widely used software packages were not FOSS, limiting the ability of the authors to share workflows with others.
MEGA was the most widely cited software in this evaluation reflecting the widespread use of sequence analysis in the field of plant pathology.
The second, third and fourth, and eighth most frequently cited software packages were all statistical programs.
SAS remains firmly entrenched in the discipline as the choice of software for statistical analysis followed by SPSS but R, the most frequently occurring FOSS, is not far removed from SPSS in fourth place with Statistica and Genstat tied for eighth.
It is interesting to note however, that even with the popularity of Python in scientific programming it only appears in the survey twice in 2020 for both articles, not enough times to be in the top 10 most-cited.

@@

We acknowledge that the number of article that are cross-scored is low.
However, our aim was not to test inter-rater reproducibility.
Our goal was to evaluate as many articles as we realistically possible to obtain as broad an overview of the discipline as possible.
Because of this, aside from the five that were used to compare the inter-rater differences, each article was evaluated by only one of this paper's co-author.
Therefore, we concluded that the inter-rater agreement was good enough for this paper to illustrate that code and data are not being widely shared with peer-reviewed publications.

@@

Plant pathology is not alone in the lack of reproducibility and the closed-nature of data and code.

@Seibold2021 carried out a similar study on the types of methods and software used in PLOS ONE articles and whether the were able to reproduce the data analysis using open source software for papers that dealt with longitudinal data and used mixed effect models or generalized estimating equations for analysis and provided the data.
They found that most articles provided tables and visualizations only.
However, unlike our analysis, even with the data being supplied they found that only one article used open source software, whereas R prominently featured in the articles we examined along with other free open-source software packages.
Not surprisingly they found that replication was mostly difficult and required the results to be reverse engineered or contacting the authors.
They were unable to reproduce the results for for three articles and only parts of another two.
They needed to contact the articles' authors for all articles save two.
Their main conclusion was that reproducing papers is difficult if no code is provided and puts an undue burden on those interested in reproducing the work.

Other disciplines related to plant pathology have published guides or started communities to promote and support tools and methods that promote good practices that lead to open and reproducible research.In entomology @Wittman2020 published "A Guide and Toolbox to Replicability and Open Science in Entomology", which documented both the benefits and drawbacks of an open and reproducible workflow.
Weed scientists started a community to promote openness in the discipline, Open Weed Science, which stated that they valued, "open access knowledge-sharing as a strategy to enhance reproducibility within our discipline", but the website is no longer available and the Twitter account, \@OpenWeedSci, appears to be inactive when viewing the account as of October 2022 [@OpenWeedScience2021].
@SerradaCruz2019 proposed a framework for capturing provenance of data-centric agronomic experiments using an R-based workflow, 'RFlow', which would allow reuse of R scripts using generic workflows and saving data and metadata to web repositories that other researchers and referees can browse using web interfaces.
@Fernandez2022 provide an R package, 'cndcR' [@Fernandez2022b], that supports the manuscript: "Dataset characteristics for the determination of critical nitrogen dilution curves: from the past to new guidelines", [@Fernandez2022a].
The package contains source data and R code that @Fernandez2022a used for their sensitivity analyses when fitting critical N dilution curves in crop species using Bayesian models. 
The package allows the user to recreate the figures in the manuscript and includes the all of the data files required to produce all results and figures presented in the manuscript.
More broadly within agricultural research, articles have been published highlighting the advantages of, the need for and how to publish Findable, Accessible, Interoperable and Reusable (FAIR) data in agriculture [@Smith2018; @Arnaud2020; @Stacey2022].
However, we were unable to find any quantification of the reproducibility of any of these related disciplines as we have done here.

@@

The first layer is including tables of raw data or code with the paper as a supplemental file or even within the paper as tables where possible.
This is suitable for studies that may have a small data set or simple analysis or for demonstrations purposes as @Madden2015 demonstrate in their discussion regarding the use of $P$ values in statistical analyses where they supply an *e-**X**tra* with reproducible examples for readers to refer to.
@Hill2019 also shared code and example data to reproduce the results of the paper as supplementary materials, citing all packages used and the versions.
The scripts provided used an R package, *checkpoint* [@Ooi2022], to provide a mechanism to help ensure reproducibility by installing the package versions used by the authors that were necessary to reproduce the work rather than defaulting to the latest versions, which may cause failures due to changes in the package code-bases.
However, the scripts did not run unhindered as the supplementary materials suggested without changes to the data.
But with some modifications to the data, and file name changes, the scripts ran allowing examples of the research to be reproduced using the definition of @Peng2009.
One drawback to this approach is that journals are often not equipped to handle code, e.g., script files, that may be developed as a part of the research process and so they are often archived as PDF or Word documents which hinder the ability to easily ingest and start working with them or if they are provided in a native text format, they do not render via the web properly.
Though this should be possible given some effort from the publishing journal to share simple text files rather than binary formats along with the proper instructions and handling of these file formats rather than a one-size-fits all approach that we commonly see for supplementary materials.
For example, the R scripts for @Hill2019 were provided as supplementary materials but do not actually appear in the browser window when requested, which may confuse readers and makes downloading them more difficult.
While this allows the reader to quickly view the extra materials and a DOI is assigned as a part of the article itself, the data or code are not readily findable and accessible through a searchable database.
Furthermore, in many cases this does not allow prompt access to the data and running the code because of a journal paywall.

@@

Taking this approach even further, packaging the full analysis in a containerized software application (Docker) is a way to help ensure computational reproducibility but at the expense of added complexity [@Merkel2014].
Docker is an open-source containerization platform that enables users to package several applications and an operating system into containers, thereby standardizing the executable components by combining application source code (or analyses) with the operating system required to run that analysis on a user's computer.
However, there are drawbacks to using Docker.
It can be difficult to understand for a new user and new platforms, like the Apple M1 chip, may not be fully supported, which hinder the ability to share the container.
But, in most cases using an open-source language like R, Julia or Python allows us to share our work in a fashion where we know that the analysis will run exactly the same on every computer.
For more on using Docker for reproducible research, we would refer readers to @Nuest2020.
As an example of this approach, @Khaliq2020 provided a research compendium as a Docker container with a DOI and a full R package.
This enables readers to fully replicate their analysis of *Ascochyta rabiei* conidia dispersal in chickpea using the data collected.
It also allows for more in-depth investigation by stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this layer is employed, tools such as Binder, <https://mybinder.org>, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks, a "web application for creating and sharing computational documents" [@jupyter2022], as Mioroni et al. used [-@Miorini2018; -@Miorini2019].
@Kamvar2015 took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

These layers provide increasing openness and reproducibility so Layer 1 should not be viewed as the starting point before moving on to Layer 2 and then Layer 3.
Rather the layers are provided to give the reader some idea of how much benefit end-users of the computational materials and data will derive at the expense of the increasing complexity.
Authors need not start with Level 1 before moving to Level 2 and then Level 3.
If they have the ability to start with depositing the computational materials and data and generating DOIs all parties will benefit from it.
However, we recognize that these steps may not be appropriate or attractive for various reasons and so have detailed other layers that improve openness and reproducibility.

>>>>

However, in the biological sciences it is not always possible to use identical test material or environmental conditions, e.g., field trials that span years and locations or complex glasshouse experiments.
In other cases, there is insufficient time and resources to reproduce the entire study.
Therefore, we will follow Peng's [-@Peng2009] definition that provides clear guidelines for a minimum standard of "reproducible research":

@@

The criteria scores were checked for inconsistencies using the evaluators' comments in the spreadsheet and based on the software used.
In some cases, these comments were used to adjust the criteria score to align with the scoring definitions.
This was not possible in all cases but only in cases where the evaluator had left comments that could be checked and acted upon.
For example, a reviewer noted that all data were deposited in GenBank, but the data score was '0', this was changed to '3' to align with the definition.
Or in other cases, some software packages used were Excel add-ins, but Excel was not recorded as being used; Excel was added to the list of software used to rectify this.

@@

Much of the software that was used in the papers examined were Free Open-Source Software (FOSS) packages, which means that the workflow can easily be recreated by anyone with the proper skills in using the software.
However, the top three most widely used software packages were not FOSS, limiting the ability of the authors to share workflows with others.
MEGA was the most widely cited software in this evaluation reflecting the widespread use of sequence analysis in the field of plant pathology.
The second, third and fourth, and eighth most frequently cited software packages were all statistical programs.
SAS remains firmly entrenched in the discipline as the choice of software for statistical analysis followed by SPSS but R, the most frequently occurring FOSS, is not far removed from SPSS in fourth place with Statistica and Genstat tied for eighth.
It is interesting to note however, that even with the popularity of Python in scientific programming it only appears in the survey twice in 2020 for both articles, not enough times to be in the top 10 most cited.

@@

We acknowledge that the number of articles that are cross scored is low.
However, our aim was not to test inter-rater reproducibility.
Our goal was to evaluate as many articles as we realistically possible to obtain as broad an overview of the discipline as possible.
Because of this, aside from the five that were used to compare the inter-rater differences, each article was evaluated by only one of this paper's co-author.
Therefore, we concluded that the inter-rater agreement was good enough for this paper to illustrate that code and data are not being widely shared with peer-reviewed publications.

@@

Plant pathology is not alone in the lack of reproducibility and the closed nature of data and code.

@Seibold2021 carried out a similar study on the types of methods and software used in PLOS ONE articles and whether they were able to reproduce the data analysis using open-source software for papers that dealt with longitudinal data and used mixed effect models or generalized estimating equations for analysis and provided the data.
They found that most articles provided tables and visualizations only.
However, unlike our analysis, even with the data being supplied they found that only one article used open-source software, whereas R prominently featured in the articles we examined along with other free open-source software packages.
Not surprisingly they found that replication was mostly difficult and required the results to be reverse engineered or contacting the authors.
They were unable to reproduce the results for for three articles and only parts of another two.
They needed to contact the articles' authors for all articles save two.
Their main conclusion was that reproducing papers is difficult if no code is provided and puts an undue burden on those interested in reproducing the work.

Other disciplines related to plant pathology have published guides or started communities to promote and support tools and methods that promote good practices that lead to open and reproducible research.In entomology @Wittman2020 published "A Guide and Toolbox to Replicability and Open Science in Entomology", which documented both the benefits and drawbacks of an open and reproducible workflow.
Weed scientists started a community to promote openness in the discipline, Open Weed Science, which stated that they valued, "open access knowledge-sharing as a strategy to enhance reproducibility within our discipline", but the website is no longer available and the Twitter account, \@OpenWeedSci, appears to be inactive when viewing the account as of October 2022 [@OpenWeedScience2021].
@SerradaCruz2019 proposed a framework for capturing provenance of data-centric agronomic experiments using an R-based workflow, 'RFlow', which would allow reuse of R scripts using generic workflows and saving data and metadata to web repositories that other researchers and referees can browse using web interfaces.
@Fernandez2022 provide an R package, 'cndcR' [@Fernandez2022b], that supports the manuscript: "Dataset characteristics for the determination of critical nitrogen dilution curves: from the past to new guidelines", [@Fernandez2022a].
The package contains source data and R code that @Fernandez2022a used for their sensitivity analyses when fitting critical N dilution curves in crop species using Bayesian models. 
The package allows the user to recreate the figures in the manuscript and includes all of the data files required to produce all results and figures presented in the manuscript.
More broadly within agricultural research, articles have been published highlighting the advantages of, the need for and how to publish Findable, Accessible, Interoperable and Reusable (FAIR) data in agriculture [@Smith2018; @Arnaud2020; @Stacey2022].
However, we were unable to find any quantification of the reproducibility of any of these related disciplines as we have done here.

@@

The first layer is including tables of raw data or code with the paper as a supplemental file or even within the paper as tables where possible.
This is suitable for studies that may have a small data set or simple analysis or for demonstrations purposes as @Madden2015 demonstrate in their discussion regarding the use of $P$ values in statistical analyses where they supply an *e-**X**tra* with reproducible examples for readers to refer to.
@Hill2019 also shared code and example data to reproduce the results of the paper as supplementary materials, citing all packages used and the versions.
The scripts provided used an R package, *checkpoint* [@Ooi2022], to provide a mechanism to help ensure reproducibility by installing the package versions used by the authors that were necessary to reproduce the work rather than defaulting to the latest versions, which may cause failures due to changes in the package codebases.
However, the scripts did not run unhindered as the supplementary materials suggested without changes to the data.
But with some modifications to the data, and file name changes, the scripts ran allowing examples of the research to be reproduced using the definition of @Peng2009.
One drawback to this approach is that journals are often not equipped to handle code, e.g., script files, that may be developed as a part of the research process and so they are often archived as PDF or Word documents which hinder the ability to easily ingest and start working with them or if they are provided in a native text format, they do not render via the web properly.
Though this should be possible given some effort from the publishing journal to share simple text files rather than binary formats along with the proper instructions and handling of these file formats rather than a one-size-fits all approach that we commonly see for supplementary materials.
For example, the R scripts for @Hill2019 were provided as supplementary materials but do not actually appear in the browser window when requested, which may confuse readers and makes downloading them more difficult.
While this allows the reader to quickly view the extra materials and a DOI is assigned as a part of the article itself, the data or code are not readily findable and accessible through a searchable database.
Furthermore, in many cases this does not allow prompt access to the data and running the code because of a journal paywall.

@@

Taking this approach even further, packaging the full analysis in a containerized software application (Docker) is a way to help ensure computational reproducibility but at the expense of added complexity [@Merkel2014].
Docker is an open-source containerization platform that enables users to package several applications and an operating system into containers, thereby standardizing the executable components by combining application source code (or analyses) with the operating system required to run that analysis on a user's computer.
However, there are drawbacks to using Docker.
It can be difficult to understand for a new user and new platforms, like the Apple M1 chip, may not be fully supported, which hinder the ability to share the container.
But, in most cases using an open-source language like R, Julia or Python allows us to share our work in a fashion where we know that the analysis will run the same on every computer.
For more on using Docker for reproducible research, we would refer readers to @Nuest2020.
As an example of this approach, @Khaliq2020 provided a research compendium as a Docker container with a DOI and a full R package.
This enables readers to fully replicate their analysis of *Ascochyta rabiei* conidia dispersal in chickpea using the data collected.
It also allows for more in-depth investigation by stepping through other points where weather data were investigated, and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this layer is employed, tools such as Binder, <https://mybinder.org>, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks, a "web application for creating and sharing computational documents" [@jupyter2022], as Mioroni et al. used [-@Miorini2018; -@Miorini2019].
@Kamvar2015 took a slightly different approach by including all files necessary for the analysis and most output files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

These layers provide increasing openness and reproducibility so Layer 1 should not be viewed as the starting point before moving on to Layer 2 and then Layer 3.
Rather the layers are provided to give the reader some idea of how much benefit end-users of the computational materials and data will derive at the expense of the increasing complexity.
Authors need not start with Level 1 before moving to Level 2 and then Level 3.
If they can start with depositing the computational materials and data and generating DOIs all parties will benefit from it.
However, we recognize that these steps may not be appropriate or attractive for various reasons and so have detailed other layers that improve openness and reproducibility.
